Global patterns in forest productivity

### 2019-09-08 manuscript/manuscript draft.Rmd
quick pass through entire manuscript
see issues #21 and #25 for more.

*Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig. 2). Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 3).

@@

@@

@@

*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature, with temperature-related climate variables coming out as strong predictors of productivity. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET (fig. 5). We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig. 5). ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

In contrast, there is no significant effect of precipitation seasonality on productivity.

@@

Our results show that productivity decreases linearly with latitude. Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity. While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables. Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP. There is a significant interaction between MAT and MAP.  We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.

@@

>>>>

In total, we analyzed ## records from ## C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years from # distinct geographic areas (Fig. 1, Table 2). [**Add a bit about geographic representation**]

@@

*How do carbon fluxes vary with latitude?*

All major carbon fluxes increase linearly with decreasing latitude (fig. 3).

@@

[*Relationships and differences among fluxes- *] In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig. 2). That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

[**move differences among R2s here?**]

@@

*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig. 5). ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

In contrast, there was no significant effect of precipitation seasonality on productivity.

@@

Our results show that productivity decreases linearly with latitude. Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale. While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically. Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP. There is a significant interaction between MAT and MAP.  We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.

@@

**[needs strong concluding paragraph]**
